---
title: "Practival Machine Learning Course Project"
author: "Vasily Vinogradov"
date: "November 21, 2015"
output: html_document
---

#Getting and cleaning data

```{r, echo=TRUE}
library(caret)
library(knitr)
trn <- read.csv("pml-training.csv", na.strings=c("NA",""), strip.white=T)
tst <- read.csv('pml-testing.csv', na.strings=c("NA",""), strip.white=T)
```

We can see that data contains many NA values that means that not all features will be very useful for prediction. Variables that contains "kurtosis","skewness","max","min","amplitude","avg","stddev","var" in names looks like aggregate results for previous rows. Therefore, we cannot use them as features.

```{r,echo=TRUE}
column_patterns <- c("kurtosis","skewness","max","min","amplitude","avg","stddev","var")
columns_to_exclude <- c(1:7,unlist(sapply(column_patterns, function(pat){grep(pat, names(trn), value = F)})))

trains <- trn[, -columns_to_exclude]
tests <- tst[, -columns_to_exclude]
dim(trains)
```

We have 52 features left (53 with classe) to build our model.


#Processing data

Lets split our training dataset into two: training and validation. 

```{r, echo=TRUE,warning=FALSE}
set.seed(1024)
inTrain <- createDataPartition(trains$classe, p=0.7, list=F)
training_set <- trains[inTrain,]
validation_set <- trains[-inTrain,]
```

#Training

Now we can train a Random Forest model on the training_set. We choose parameters to reduce running time.

```{r, echo=TRUE,warning=FALSE}
ctrl<-trainControl(method = "cv", number = 5,allowParallel = TRUE)
model_rf<-train(classe ~ ., data=training_set, model="rf", trControl=ctrl)
```

#Validation

To validate our model we perfomr prediction on validation_set.

```{r, echo=TRUE,warning=FALSE}
prediction_rf<-predict(model_rf, newdata=validation_set)
```

Now we can compare predictions and known classes and find out-of-model error from confusionMatrix function

```{r,warning=FALSE}
cm<-confusionMatrix(prediction_rf, validation_set$classe)
cm
```

out-of-model error is 

```{r}
error<-unname(1-cm$overall["Accuracy"])
error
```


#Prediction

Let us do prediction with our trained model on unknown set.

```{r}
predict_final<-predict(model_rf,newdata=tests)
predict_final
```


